Unrecognized States: Theory and Cases

### 2017-06-30 editedvolume.tex
fix minor typos in text; add notes from initial read

However, despite past failures, or theory suggests that a sufficiently motivated patron can induce negotiated settlement if it so chooses. The means through which the international community might induce negotiated settlement are discussed in detail in the section on policy implications. It is notable, however, that we do not expect any future cases of peacefully negotiated independence.  While negotiated reunification is the preferred outcome of the international community, and they may be willing spend to achieve it, recognized statehood is generally not the preferred outcome of the patron or the international community. Our analysis suggests that the most likely path to recognized statehood is, and will remain, military defeat of the home state.

@@

Gagauzia achieved de facto independence at the time of the Soviet Union's collapse, but agreed to rejoin Moldova in 1994 as an autonomous region.  While Gagauzia was granted substantial autonomy under the Moldovan Law on the Special Legal Status of Gagauzia, when the governor of Gagauzia, Dmitrii Croiter, moved to assert these powers in 1999, the Moldovan government balked.  By 2002, Croiter was forced to resign, effectively deposed by the Moldovan government.  The Moldovan government jailed a number of other Gagauz politicians, and while Gagauz autonomy was enshrined in the Moldovan constitution in 2003, the de facto level of autonomy has been limited by continued by central government meddling in less-than-free regional elections.\footnote{Protsyk (2010) provides an account of the "salami tactics" by which Moldovan authorities have gradually reclaimed powers originally granted to the regional government.} The payoffs to Gagauzia for ceding have turned out to be quite low, and a similar fate can rationally be expected by other unrecognized states who choose to cede.\footnote{Roper (2002) argues that secessionists in Transnistria are wary of negotiated reunification precisely because of the creeping re-centralization they have observed in Gagauzia.}

@@

In Southern Sudan, the international community invested substantial resources to help negotiate a settlement and to ensure that the Sudanese government government both allowed the promised a referendum and respected its results. While the international community acted in Southern Sudan to enforce independence, not autonomy, it has shown itself capable of enforcing difficult concessions by the home state government. This bodes well for the future credibility of the international community as a third-party enforcer.  However, the role of the international community in enforcing other past agreements might give secessionists pause. For example, a referendum on independence in Western Sahara, which the UN ruled to be necessary more than thirty years ago, has never come to pass.\footnote{For a thorough analysis of the Western Sahara case, see Zunes and Mundy (2010).}  Nonetheless, it is possible for the international community to invest resources to enforce agreements, allowing for negotiated settlements that would otherwise be impossible to reach.

>>>>

However, despite past failures, our theory suggests that a sufficiently motivated patron can induce negotiated settlement if it so chooses. The means through which the international community might induce negotiated settlement are discussed in detail in the section on policy implications. It is notable, however, that we do not expect any future cases of peacefully negotiated independence.  While negotiated reunification is the preferred outcome of the international community, and they may be willing spend to achieve it, recognized statehood is generally not the preferred outcome of the patron or the international community. Our analysis suggests that the most likely path to recognized statehood is, and will remain, military defeat of the home state.

@@

Gagauzia achieved de facto independence at the time of the Soviet Union's collapse, but agreed to rejoin Moldova in 1994 as an autonomous region.  While Gagauzia was granted substantial autonomy under the Moldovan Law on the Special Legal Status of Gagauzia, when the governor of Gagauzia, Dmitrii Croiter, moved to assert these powers in 1999, the Moldovan government balked.  By 2002, Croiter was forced to resign, effectively deposed by the Moldovan government.  The Moldovan government jailed a number of other Gagauz politicians, and while Gagauz autonomy was enshrined in the Moldovan constitution in 2003, the de facto level of autonomy has been limited by continued central government meddling in less-than-free regional elections.\footnote{Protsyk (2010) provides an account of the "salami tactics" by which Moldovan authorities have gradually reclaimed powers originally granted to the regional government.} The payoffs to Gagauzia for ceding have turned out to be quite low, and a similar fate can rationally be expected by other unrecognized states who choose to cede.\footnote{Roper (2002) argues that secessionists in Transnistria are wary of negotiated reunification precisely because of the creeping re-centralization they have observed in Gagauzia.}

@@

In Southern Sudan, the international community invested substantial resources to help negotiate a settlement and to ensure that the Sudanese government both allowed the promised a referendum and respected its results. While the international community acted in Southern Sudan to enforce independence, not autonomy, it has shown itself capable of enforcing difficult concessions by the home state government. This bodes well for the future credibility of the international community as a third-party enforcer.  However, the role of the international community in enforcing other past agreements might give secessionists pause. For example, a referendum on independence in Western Sahara, which the UN ruled to be necessary more than thirty years ago, has never come to pass.\footnote{For a thorough analysis of the Western Sahara case, see Zunes and Mundy (2010).}  Nonetheless, it is possible for the international community to invest resources to enforce agreements, allowing for negotiated settlements that would otherwise be impossible to reach.
